---
title: "Assignment 7 By Team 1"
author: "Odianosen Akhibi, Shun-Lung Chang, Juliana Nair"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center', fig.height = 2.5)
```

```{r library packages, include=FALSE}
library(magrittr)
library(ggplot2)
library(knitr)
```

This study was conducted in R, and the source code can be found [here](https://github.com/slchangtw/data_analytics_projects/blob/master/assignment_7.Rmd).

# 1. Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables.

```{r}
# set the random number seed
set.seed(45)

normal <- rnorm(1000, mean = 100, sd = 1) %>% 
    matrix(20, 50) %>%
    data.frame()
uniform <- runif(1000, min = -100, max = -50) %>% 
    matrix(20, 50) %>%
    data.frame()
exponential <- rexp(1000, rate = 1) %>% 
    matrix(20, 50) %>%
    data.frame()

dat <- rbind(normal, uniform, exponential)

dim(dat)
```

# 2. Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. 

```{r}
pca <- prcomp(dat) 
```

```{r, echo=FALSE}
two_pcs <- data.frame(index = as.factor(1:60), 
                           PC1 = pca$x[, 1], 
                           PC2 = pca$x[, 2],
                           group = c(rep("Normal", 20), rep("Uniform", 20), rep("Exponential", 20)))
```

```{r, echo=FALSE}
ggplot(two_pcs) +
    geom_point(aes(x = index, y = PC1, color = group)) + 
    theme_bw(base_size = 9) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r, echo=FALSE}
ggplot(two_pcs) +
    geom_point(aes(x = index, y = PC2, color = group)) + 
    theme_bw(base_size = 9) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# 3. Perform K-means clustering of the observations with K = 3. How well do the clusters that you obtained in K-means clustering compare to the true class labels?

```{r}
kc_3 <- kmeans(dat, center = 3)

table(kc_3$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```

# 4. Perform K-means clustering with K = 2. Describe your results.

```{r}
kc_2 <- kmeans(dat, center = 2)

table(kc_2$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```

# 5. Now perform K-means clustering with K = 4, and describe your results.

```{r}
kc_4 <- kmeans(dat, center = 4)

table(kc_4$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```

# 6. Now perform K-means clustering with K = 3 on the first two principal component score vectors, rather than on the raw data. That is, pserform K-means clustering on the 60 × 2 matrix of which the first column is the first principal component score vector, and the second column is the second principal component score vector. Comment on the results.

```{r}
kc_pc_3 <- kmeans(data.frame(two_pcs$PC1, two_pcs$PC2), centers = 3)

table(kc_pc_3$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```

# 7. Using the `scale()` function, perform K-means clustering with K = 3 on the data after scaling each variable to have standard deviation one. How do these results compare to those obtained in (3)? Explain.

```{r}
dat_scaled <- scale(dat)
kc_3 <- kmeans(dat_scaled, center = 3)

table(kc_3$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```

# 8. Use the scaled variables and run a PCA on them. Now perform K-means clustering with K = 3 on the first two principal component score vectors, rather than on the raw data. How do these results compare to those obtained in (3) and (7)? Explain.

```{r}
pca_scaled <- prcomp(dat, center = TRUE, scale. = TRUE) 

kc_pc_scaled_3 <- kmeans(data.frame(pca_scaled$x[, 1], pca_scaled$x[, 2]), centers = 3)

table(kc_pc_scaled_3$cluster, c(rep(1, 20), rep(2, 20), rep(3, 20))) %>% kable()
```