---
title: "Assignment 6 By Team 2"
author: "Shun-Lung Chang, Deepika Ganesan, Deepankar Upadhyay"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. First, perform hierarchical clustering on the states.

## a) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states
```{r}
hc <- hclust(dist(USArrests))
hc
```
## b) Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

```{r, echo=FALSE}
plot(hc)
```


## c) Hierarchically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one

```{r}
scaled_USArrests <- scale(USArrests) 
hc_scaled <- hclust(dist(scaled_USArrests))
plot(hc_scaled)
```

## d) What effect does scaling the variables have on the hierarchical clustering obtained? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed? Provide a justification for your answer

# 2. Perform k-means clustering, selecting a suitable range for k. Compare the results with the ones from question 1

```{r} 
set.seed(42)
k_max <- 10
wss <- sapply(3:k_max,  function(k) { kmeans(scaled_USArrests, k, iter.max = 50)$tot.withinss})

plot(3:k_max, wss,
     type = "b", pch = 19, frame = FALSE, 
     xlab = "Number of clusters K",
     ylab = "Total within-clusters sum of squares")
```